### 2020-12-01 paper/localpsf.tex
editing intro

In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, which have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
\begin{equation}
\label{eq:kernel_representation}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
\begin{description}
\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$. Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. 
\item[Non-negative:] $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
\item[Local support:] $A(y, x)$ is supported near $y=x$.
\item[Local translation invariance:] $A(y+h, x+h) \approx A(y,x)$
if $h$ is not too large.
\end{description}
If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of a convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. The convolution kernel,
\begin{equation*}
	\varphi_x(y) := A(y+x,x)
\end{equation*}
is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$. Locally translation invariant operators therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.

>>>>

In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
\begin{equation}
\label{eq:kernel_representation}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
\begin{description}
\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$. Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. 
\item[Non-negative:] $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
\item[Local support:] $A(y, x)$ is supported near $y=x$.
\item[Local translation invariance:] $A(y+h, x+h) \approx A(y,x)$
if $h$ is not too large.
\end{description}

If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of the convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$. The convolution kernel,
\begin{equation*}
	\varphi_x(y) := A(y+x,x),
\end{equation*}
is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. We therefore sample $\varphi_{x_i}$ at a collection of points $\{x_i\}_{i=1}^r \subset \Omega$, and patch together the local convolution operators $u \mapsto \varphi_{x_i} \ast u$ using smooth weighting functions $w_i(x)$. This yields a product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
\begin{equation*}
	\left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)}.
\end{equation*}
Here $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$.

While kernel entries $\mathcal{A}(y,x)$ are not easily computable, kernel entries $\widetilde{A}(y,x)$ are given by the formula
\begin{equation*}
	\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
\end{equation*}
